---
title: "Module 3"
author: "Orhun Kok"
date: "9/25/2020"
output: html_document
---

### Question 1.

The answer is B because a range is given, and therefore it is a continous random variable. This narrows the answers down to the integrals. From here, we see that the function is for x > 1 and given the range of 0 < x < 4, the only choice that fits all of that is B. 

### Question 2.
```{r Question 2, echo = FALSE}
# P(X = 1) - Simple subsitution of 1 for variable x
a <- (2^1/(factorial(1))*exp(-2))
paste(a,"is P(X = 1)")
# P(-2<X<4) - Does not include P(4) and lower limit can't go to -2 because x is a defined list
b1 <- (2^0/(factorial(0))*exp(-2))
b2 <- (2^1/(factorial(1))*exp(-2))
b3 <- (2^2/(factorial(2))*exp(-2))
b4 <- (2^3/(factorial(3))*exp(-2))
b <- sum(b1,b2,b3,b4)
paste(b, " is P(-2<X<4)")
```

### Question 3.

The n value in a binomial represents the number of children in this case, and the p value represents the probability of albinism. Therefore, n = 3 and p = 0.25.

### Question 4.
```{r Question 4, echo = FALSE}
# P(Y<= 2) = P(Y = 0) P(Y = 1) + P(Y = 2)
set.Y <- c(0,1,2)
c <- sum(dbinom(set.Y, size = 3, p = 0.25))
paste(c, "is P(Y <= 2)")
#E(Y) = np
paste(3*0.25, "is E(Y)")
#Var(Y) = np*(1-p)
paste((3*0.25)*(1-0.25), "is Var(Y)")
```
### Question 5.
```{r Question 5, echo = FALSE}
# P(1 < X < 4)
X <- integrate(function(X) dchisq(X, df = 3), lower = 1, upper = 4)
paste(X[1],"with absolute error",X[2], "is P(1<X<4)")
# E(X) equals the degrees of freedom for chi-square
paste(3,"is E(X)")
# Var(X) equals twice the degrees of freedom for chi-square
paste(6,"is Var(X)")
# Monte Carlo Simulation with sample size of 100,000 and comparing with original
Monte <- rchisq(n = 100000, df = 3)
Monte <- mean((1 < Monte) & (Monte < 4))
paste(Monte,"is the Monte Carlo estimate, compared to the", X[1], "answer from above")
```

### Question 6.
```{r Question 6, echo = FALSE}
# E(Y) = 4(E(X))-10 according to linear transformation rules
meanY <- (4*5) - 10
paste(meanY,"is E(Y)")
# Var(Y) = Var(4X-10) = a^2(Var((X))) = 16(Var(X))
varY <- 16*10
paste(varY,"is Var(Y)")
# Degree of freedom = mean, Variance = 2 * degrees of freedom = 20. Therefore, 20 != varY
print("Y does not follow a chi square distrubution with df = 10 because while the mean does equal 10, the variances do not match")
```

### Question 7.
```{r Question 7, echo = FALSE}
# P(1<X<1.6) = P(X>1.6) - P(X<1)
d <- pnorm(1.6,mean = 1.6, sd = 0.4) - pnorm(1,mean = 1.6, sd = 0.4)
paste(d,"is the probability that a randomly chosen patient has Zyxin gene expression values between 1 and 1.6")
# Monte Carlo Simulation with sample size of 500,000
Monte2 <- rnorm(n = 500000, mean = 1, sd = 0.4)
Monte2 <- mean((1 < Monte2) & (Monte2 < 1.6))
paste(Monte2,"is the Monte Carlo estimate")
# Exactly 2 out of 5 patients having Zyxin expression between 1 and 1.6
e <- dbinom(2,size = 5, prob = d)
paste(e," is the probability that exactly 2 out of 5 patients have Zyxin expression between 1 and 1.6 ")
```

### Question 8.
```{r Question 8, echo = FALSE}
# Integration of X~F and Y~F with lower limit at 0 and upper at Inf as per instructions for mean calculation
integralmeanXF <- integrate(function(x) x*df(x,df1 = 2, df2 = 5), lower = 0, upper = Inf)
integralmeanYF <- integrate(function(y) y*df(y,df1 = 10, df2 = 5), lower = 0, upper = Inf)
# Integration of X~F and Y~F with lower limit -Inf and upper at Inf for variance calculation
integralvarianceXF <- integrate((function(c) ((c - integralmeanXF$value)^2)*df(c,df1 = 2, df2 = 5)), lower = -Inf, upper = Inf)
integralvarianceYF <- integrate((function(k) ((k - integralmeanYF$value)^2)*df(k,df1 = 10, df2 = 5)), lower = -Inf, upper = Inf)
# Mean = n/n-2 | Variance = (2n^2(m+n-2))/(m(n-2)^2)(n-4) for n > 4. Formulas found in table 3.4.1
meanXF <- 5/(5-2)
varianceXF <- (((2*((5)^2)))*(2+5-2))/((2*(5-2)^2)*(5-4))
paste(meanXF, "is the mean of X~F(m=2,n=5) and", varianceXF, "is the variance of X~F(m=2,n=5)")
meanYF <- 5/(5-2)
varianceYF <- (((2*((5)^2)))*(10+5-2))/((10*(5-2)^2)*(5-4))
paste(meanYF, "is the mean of Y~F(m=10,n=5) and", varianceYF, "is the variance of Y~F(m=10,n=5)")
# Values from integration when compared to manual calculation are the same
paste(integralmeanXF$value,"and",integralmeanYF$value,"are the mean values from part A for X~F and Y~F respectively, and", integralvarianceXF$value,"and", integralvarianceYF$value, "are the variance values from part A for X~F and Y~F respectively. When comparing the values from Part A and B, you can see they match.")
```
